---
title: "Tutorial on R and R Studio (Part III)"
author: "Le Wang"
output:
  html_document:
    number_sections: true
    toc: true
    toc_depth: 3    
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In class, you can download this set of code [here](lecture/intro-to-R_part3.R)


# Data types in `R`

Here we discuss various types of data in `R`. `R` works with numerous data types. Some of the most basic types to get started are:

	1. **Decimal values** like 4.5 are called numerics.
	2. **Natural numbers** like 4 are called integers. Integers are also numerics.
	3. **Boolean values** (TRUE or FALSE) are called logical.
	4. **Text** (or string) values are called characters.

Note how the quotation marks on the right indicate that "some text" is a character. You can check the type of data by using `class()`. For example, we had the variable x above

```{r}
  x <- "Melody to Funkytown!"
  class(x)
```



# A collection of data points

## Vectors  	

### Some basic examples 

```{r}
v <- c(1, 2, 3, 4)		
v
v <- 1:4		
v
v <- seq(from = 0, to = 0.5, by = 0.1)		
v


#Vectors with strings 	
v_colors <- c("blue", "yellow", "light green")	
v_colors	

```

Here `c()` stands for **concatenate**


### Vector selection (Indexing/reassigning elements)

```{r}
v_colors[2]         
v_colors[c(1, 3)]   

v_colors[2:3]  <- c("red", "purple")	
v_colors 	

```

See [Datacamp course](https://campus.datacamp.com/courses/free-introduction-to-r) for other ways to subset a vector


### Names of a vector

```{r}
names(v_colors) <- c("Le", "Austin", "Heather")
```


In the [Datacamp course](https://campus.datacamp.com/courses/free-introduction-to-r) you will learn other operations of vectors.


## Matrix

### create a new matrix
```{r}
matrix<-matrix(1:4, byrow = TRUE, nrow = 2)
matrix
```

Note that `:` means every number from 1 to 4. In the `matrix()` function:

    1. The first argument is the collection of elements that `R` will arrange into the rows and columns of the matrix. Here, we use 1:4 which is a shortcut for c(1, 2, 3, 4).
    2. The argument `byrow` indicates that the matrix is filled by the rows. If we want the matrix to be filled by the columns, we just place `byrow = FALSE`.
    3. The third argument `nrow` indicates that the matrix should have three rows.	

### rename a matrix

```{r}
	rownames(matrix) <- c("Le", "Bingyue")
	colnames(matrix) <- c("Yes", "No")
	
	matrix
```

	
## Dataframe

### What's a data frame?

Our data often contain various types of data or variables. For example, in the CPS data that you will 
be dealing with, you will have information on income (numeric), gender (logical), race (string)
A data frame can have many types of data, and each variable is treated as columns and 
the observations (or units) as rows. 


### Why use data.frames?

    1. data.frames - unlike general arrays or, specifically, matrices in R - can hold variables of different flavors (heuristic term defined later), such as character data (subject ID or name), quantitative data (white blood cell count), and categorical information (treated vs. untreated)
    2. data.frames package related variables neatly together,keeping them in sync vis-a-vis row order applying any filtering of observations uniformly
    3. Most functions for inference, modelling, and graphing are happy to be passed a data.frame via a data =?argument as the place to find the variables you're working on; the latest and greatest packages actually?require?that your data be in a data.frame



### Create a data frame

Data frame with Harry Potter characters

```{r}
	name <- c("Harry", "Ron", "Hermione", "Hagrid", "Voldemort")	
	height <- c(176, 175, 167, 230, 180)	
	gpa <- c(3.4, 2.8, 4.0, 2.2, 3.4)	
	df_students <- data.frame(name, height, gpa)  		
	df_students	
```


Alternative way of creating DF

```{r}
	df_students <- data.frame(name = c("Harry", "Ron", "Hermione", "Hagrid",
					   "Voldemort"), 	
				  height = c(176, 175, 167, 230, 180), 	
				  gpa = c(3.4, 2.8, 4.0, 2.2, 3.4))	
	df_students

```


	
	
### Adding variable

```{r}
	df_students$good <- c(1, 1, 1, 1, 0)   	
	df_students	
```


### Features of the DF

```{r}
	dim(df_students)  		
	df_students[2, 3]               #Ron's GPA		
	df_students$gpa[2]              #Ron's GPA	

	df_students[5, ]                #get row 5
	df_students[3:5, ]              #get rows 3-5

	df_students[, 2]                #get column 2 (height)	
	df_students$height              #get column 2 (height) 
	df_students[, 1:3]              #get columns 1-3 


	df_students[4, 2] <- 255        #reassign Hagrid's height	
	df_students$height[4] <- 255    #same thing as above	
	df_students		
```

